Developmental trajectories of word recognition: a priming study in bilinguals

### 2023-11-27 manuscript/manuscript.qmd

library(knitr)
library(kableExtra)
library(ggplot2)
library(gt)
library(gtExtras)
library(patchwork)
library(english)
library(gtExtras)

@@

clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")

options(ggplot2.ordinal.fill = clrs[c(1, 4, 5)],
		ggplot2.ordinal.colour = clrs[c(1, 4, 5)],
		ggplot2.discrete.fill = clrs[c(1, 4, 5)],
		ggplot2.discrete.colour = clrs[c(1, 4, 5)],
		ggplot2.continuous.fill = ggplot2::scale_color_gradient,
		ggplot2.continuous.colour = ggplot2::scale_color_gradient)

@@

In line with previous studies in monolinguals, we further predicted that the strength of the lexical interference effects in the *Related/Non-cognate* and *Related/Cognate* conditions would be associated to participants' vocabulary size. Target word recognition should be delayed by the activation of a larger cohort of phonologically related words [@mani2011phonological; @chow2017spoken]. We defined vocabulary size as the amount of words participants were reported to understand in their dominant language by their caregivers. The choice of the dominant language for calculating vocabulary sizes is due to several reasons. First, it allows a more fair comparison between monolinguals (who do not know any language other than their dominant language) and bilinguals (who may know words in a second language). Second, since participants were tested exclusively in their dominant language, their vocabulary size in the dominant language is more likely to be associated to participants' performance in the task. Third, previous work on word recognition in bilinguals suggests that vocabulary size in the dominant language predicts participants' performance better than total vocabulary (in which vocabulary sizes in both languages are summed together) [@marchman2010vocabulary].

@@

	cols_merge(c(age_min, age_max), pattern = "{1}-{2}") |>
	cols_merge(c(age_mean, age_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(doe_catalan_mean, doe_catalan_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(doe_english_mean, doe_english_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(doe_spanish_mean, doe_spanish_sd),
			   pattern = "{1} ({2})") |>
	tab_spanner("Sample size", starts_with("n")) |>
	tab_spanner("Age (months)", matches("age")) |>
	tab_spanner("Catalan", matches("catalan")) |>
	tab_spanner("English", matches("english")) |>
	tab_spanner("Spanish", matches("spanish")) |>
	tab_spanner("Degree of Exposure (%)", matches("doe")) |>
	cols_label(n = "Testing sessions",
			   n_participants = "Participants",
			   age_mean = "M (SD)",
			   age_min = "Range",
			   doe_catalan_mean = "M (SD)",
			   doe_english_mean = "M (SD)",
			   doe_spanish_mean = "M (SD)") |>
	tab_style(cell_text(style = "italic"),
			  cells_column_labels()) |>
	tab_style(cell_text(style = "italic"),
			  cells_column_spanners())
```

@@

```{r fig-vocabulary}
#| label: fig-vocabulary-oxf
#| fig-cap: "Participant receptive vocabulary sizes across ages and language profiles. Vocabulary size scores from the recurrent participants are shown linked. "
#| fig-width: 6
#| fig-height: 4
vocab_plot_data <- vocabulary |>
	inner_join(filter(attrition_participants, is_valid_participant),
			   by = join_by(session_id)) |>
	inner_join(select(participants, location, lp, age,
					  child_id, session_id, age_group),
			   by = join_by(child_id, session_id)) |> 
	filter(is_valid_participant) |> 
	select(child_id, session_id, vocab_id, is_imputed, l1_count,
		   lp, location, age_group, age) |> 
	mutate(location = factor(location, 
							 levels = c("Oxford", "Barcelona")),
		   lp = gsub(" \\(English\\)", "", lp),
		   side = case_when(location=="Oxford" ~ "both",
		   				 lp=="Monolingual" ~ "left",
		   				 lp=="Bilingual" ~ "right"),
		   study = if_else(location=="Oxford", 
		   				"Study 1 (English)", "Study 2 (Catalan/Spanish)"))

@@

vocab_plot_data |> 
	filter(study=="Study 1 (English)") |> 
	ggplot(aes(age, l1_count, 
			   side = side)) +
	geom_point(data = filter(vocab_plot_means, 
							 study=="Study 1 (English)"), 
			   colour = clrs[1],
			   show.legend = FALSE,
			   size = 2,
			   shape = 1,
			   stroke = 1) +
	geom_smooth(method = "lm", 
				formula = "y ~ x",
				colour = clrs[1],
				fill = clrs[1]) +
	labs(x = "Age (months)",
		 y = "Vocabulary size",
		 colour = "Group",
		 fill = "Group") +
	scale_y_continuous(limits = c(0, 575),
					   breaks = seq(0, 1e3, 50)) +
	scale_x_continuous(breaks = seq(0, 50, 2)) +
	theme(panel.grid.major.y = element_line(linetype = "dotted",
											colour = "grey"),
		  legend.position = "top",
		  axis.title.x = element_blank(),
		  legend.title = element_blank())
```

@@

We created four stimuli lists, to which participants were randomly allocated. Recurrent participants were always allocated to the same list. Each list contained a different subset of prime-target-distractor triples. Target-distractor pairs were kept constant across lists: all infants, regardless of the list to which they were allocated, were presented with the *ball*-*trousers* target distractor pair. Each target-distractor pair was presented twice in their corresponding lists, alternating which word acted as target, and which as distractor. For instance, the *ball*-*trousers* pair was presented in all lists twice, once with *ball* as target, and another with *trousers* as target. Across lists, the same target-distractor pair appeared with a different prime, counterbalancing the condition to which it belonged. For instance, in list one the *ball*-*trousers* pair was preceded by the prime *bike* (*Related/Cognate*), by *butterfly* (*Related/Non-cognate*) in list two, and by *star* and *nose* (*Unrelated*) in lists three and four, respectively.

@tbl-stimuli shows a detailed summary of the stimuli properties, broken down by trial type and testing language. Trials included in each condition had equivalent length (number of phonemes) and lexical frequency. Lexical frequencies were extracted from the English corpora from the CHILDES database [@macwhinney2000childes; @sanchez2019childes] as counts per million words, and transformed into Zipf scores for easier cross-language comparison [@zipf1945meaning; @van2014subtlex]. In addition to sharing phonological onset, prime and target shared higher overall phonological similarity---as measured using the Levenshtein distance---between their phonological transcriptions---in the *Related/Cognate* and the *Related/Non-cognate* conditions, compared to the *Unrelated* condition. In *Related/Cognate* trials, the prime label in the language of test shared a higher overall phonological similarity with its translation, compared to *Related/Non-cognate* trials. Since the number of cognate and non-cognate primes was balanced in *Unrelated* trials, overall phonological similarity between prime labels prime labels in the *Unrelated* condition had an intermediate overall phonological similarity. Audio durations were also equivalent across conditions.

@@

Testing took place in a sound-proof room at the BabyLab of the University of Oxford. Participants sat on their caregivers' lap in a dimly lit testing booth while the experimenter conducted the experiment from outside. Caregivers were instructed to keep their eyes shut (to avoid recording their gaze, instead of the participant's), to be still, and to avoid interacting with the participant verbally or non-verbally. Participants sat at approximately 65 cm from the eye-tracker and a XX-in screen of $1929\times1080$ screen resolution. The study was run using a custom Matlab script, PresentMate, based on the PsychToolbox-3 extension [@kleiner2007s; @brainard1997psychophysics; @pelli1997videotoolbox]. Visual fixations were recorded using a Tobii TX 300 eye-tracker and a 23-in screen of 1920 $\times$ 1080 resolution. The Tobii Analytics SDK 3.0 was used to interact with the eye-tracking while the experiment was running. Sampling rate was set at 120 Hz. A 9-point calibration was performed before every experimental session, in which the picture of a colourful beach ball was presented. We set a 55% grey background for the screen during calibration and stimuli presentation. Auditory stimuli were presented through two loudspeakers located behind the screen, one to each side. The experimenter monitored the experimental from outside the room using a centrally located video camera place above the screen. After a successful calibration the experimenter triggered the onset of the first trial. Trials were presented uninterruptedly and without intervention of the experimenter until the 32 trials were presented, or the experimental session had to be stopped because of the participant's behaviour.

@@

After trials that matched any of the aforementioned exclusion criteria from the data set, we excluded participants who did not provide at least two valid trials in the *Cognate* condition, the *Non-cognate* condition, or the *Unrelated* condition (*n* = `r n_exc_participants`), and participants with a L1 vocabulary size lower than 42, which corresponds to 10% of the words in the OCDI vocabulary checklist (*n* = `r n_exc_participants_vocab`). The final data set included `r format(n_trials_valid, big.mark = ",")` trials from `r n_sessions_valid` testing sessions, generated by `r format(n_participants_valid, big.mark = ",")` distinct participants. Of those participants, `r n_longitudinal[1, 2]` provided data from one experimental session, `r n_longitudinal[2, 2]` provided data from two experimental sessions, and `r n_longitudinal[3, 2]` provided data from three experimental sessions. From the trials included in the final data set, `r n_trials_condition$unrelated_2$n` were *Unrelated* trials (`r n_trials_condition$unrelated$n` previously excluded), were `r n_trials_condition$non_cognate_2$n` *Related/Non-cognate* trials(`r n_trials_condition$non_cognate$n` previously excluded), and `r n_trials_condition$cognate_2$n` were *Related/Cognate* trials (`r n_trials_condition$cognate$n` previously excluded).

@@

r_version <- glue::glue(r_info$R.version$major,
						r_info$R.version$minor,
						.sep = "."
)
stan_version <- "2.33.0"

@@

model_summary <- get_posterior_summary(model_fits_oxf$fit_oxf_0,
									   data_bcn,
									   vars_dict = var_labels,
									   rope_interval = rope_interval
)

@@

```{r fig-epreds-oxf}
#| label: fig-epreds-oxf
#| fig-cap: "A) Posterior mean predictions of the time course of target fixation in the test phase. B) Posterior mean prediction of the time course of the differences in target looking time between conditions. Intervals represent the 95% CrI of the posterior predictions. Lines indicate the mean of the posterior predictions."
#| fig-width: 6.5
#| fig-height: 6.5
make_std <- function(x, mean, sd) (x - mean) / sd

@@

@@

We found strong evidence of successful word recognition across participants of all ages, but we did not observe any evidence of phonological priming. English monolingual participants from all ages showed an equivalent pattern of target looking in both Related (Related/Cognate and Related/Non-cognate) and Unrelated trials. In conclusion, we failed to replicate the original studies by @mani2010infant and @mani2011phonological. The absence of a phonological priming effect suggest that either English monolinguals did not generate implicit labels for the prime pictures presented in silence, or that, if generated, such labels did not interact with the subsequent recognition of the target word. Both candidate explanations conflict with both Mani and Plunkett’s studies, and also with previous studies suggesting that infants 12-months and older already generate internal labels when presented with pictures of familiar objects (@duta2012erp; @styles2015infant).

Adding the predictors *Age* or *Vocabulary size* as predictors in the model, in interaction with *Condition* did not increase the fit of the model. This points to neither variable having a substantial influence in participants' target looking behaviour across conditions. These results diverge from previous studies reporting an increment in word recognition speed [@fernald1998rapid; @marchman2008speed], and stronger phonological priming effects in children with larger vocabulary sizes [@mani2011phonological; @chow2017spoken]. Overall, this results suggest that our modification of the implicit naming task resulted in the loss of the originally reported effect.

@@

suppressWarnings({
	time_diffs <- participants_bcn |>
		arrange(child_id, session_id, age) |>
		mutate(n_session = 1:n()) |>
		select(child_id, session_id, session_n, age) |>
		filter(child_id %in% participants_bcn$child_id[duplicated(participants_bcn$child_id)]) |>
		pivot_wider(
			id_cols = child_id,
			names_from = session_n,
			values_from = age,
			names_repair = janitor::make_clean_names
		) |>
		rowwise() |>
		mutate(
			diff_1 = x2 - x1, diff_2 = x3 - x2,
			diff_min = min(diff_1, diff_2, na.rm = TRUE),
			diff_max = max(diff_1, diff_2, na.rm = TRUE)
		) |>
		ungroup() |>
		summarise(
			diff_min = min(diff_min, na.rm = TRUE),
			diff_max = max(diff_max, na.rm = TRUE)
		)
})
```

@@

We collected vocabulary data using parental responses to the Barcelona Vocabulary Questionnaire [BVQ, @garcia-castro2023bvq], an online vocabulary checklist developed to assess the vocabulary size of Catalan-Spanish bilingual toddlers, and inspired in several adaptations of the the Communicative Developmental Inventory [CDI, @fenson1994variability]. This questionnaire has four versions, each including a different but overlapping subset of words, from a total pool of `r n_pool` words from `r n_categories` functional-semantic categories. Each version included a Catalan and a Spanish vocabulary checklist. Catalan checklists contained between `r n_items$min[1]` and `r n_items$min[2]` words, and Spanish checklists contained between `r n_items$min[2]` and `r n_items$min[2]` words (see the Methods section of Chapter 1 for a detailed description of the questionnaire.. Participants were randomly allocated to one of the four versions. Recurrent participants were always allocated to the same version. Families received a link to the BVQ immediately after each experimental session, and were given two weeks to fill it.

@@

```{r fig-vocabulary}
#| label: fig-vocabulary-bcn
#| fig-cap: "Participant receptive vocabulary sizes across ages and language profiles. Points and intervals indicate mean and standard deviations, respectively."
#| fig-width: 6
#| fig-height: 4
vocab_plot_data |> 
	filter(study=="Study 2 (Catalan/Spanish)") |> 
	ggplot(aes(age, l1_count, 
			   colour = lp,
			   fill = lp)) +
	facet_wrap(~lp) +
	geom_point(size = 2,
			   shape = 1,
			   stroke = 1) +
	geom_smooth(method = "lm",
				formula = "y ~ x") +
	labs(x = "Age (months)",
		 y = "Vocabulary size",
		 colour = "Group",
		 fill = "Group") +
	scale_color_manual(values = clrs[c(1, 5)]) +
	scale_fill_manual(values = clrs[c(1, 5)]) +
	theme(panel.grid.major.y = element_line(linetype = "dotted",
											colour = "grey"),
		  legend.position = "top",
		  axis.title.x = element_blank(),
		  legend.title = element_blank())
```

@@

Same as in Study 1. We used a custom Matlab XXXX script using the PsychToolbox-3 extension to present the stimuli, and the Tobii Analytics SDK 3.0 to interact with the eye-tracking while the experiment was running.

@@

```{r data set}
attrition_bcn <- filter(attrition, location=="Barcelona")
n_trials <- nrow(attrition_bcn)
n_sessions <- n_distinct(attrition_bcn$session_id)
n_participants <- n_distinct(attrition_bcn$child_id)
n_trials_valid <- nrow(filter(
	attrition_bcn,
	is_valid_participant,
	is_valid_trial
))
n_sessions_valid <- sum(distinct(attrition_bcn,
								 session_id,
								 .keep_all = TRUE
)$is_valid_participant)
n_participants_valid <- n_distinct(
	attrition_bcn[attrition_bcn$is_valid_participant, ]$child_id
)

n_exc_prime <- sum(!attrition_bcn$is_valid_gaze_prime)
n_exc_test <- sum(!attrition_bcn$is_valid_gaze_test)
n_exc_test_each <- sum(!attrition_bcn$is_valid_gaze_test_each)
n_exc_participants <- sum(!distinct(attrition_bcn,
									session_id,
									.keep_all = TRUE
)$is_valid_participant)
n_exc_participants_vocab <- sum(!distinct(attrition_bcn,
										  session_id,
										  .keep_all = TRUE
)$is_valid_vocab_size)

n_longitudinal <- attrition_bcn |>
	filter(
		is_valid_participant,
		is_valid_trial
	) |>
	distinct(child_id, session_n) |>
	count(child_id, name = "times") |>
	count(times)
```

@@

We modeled the data following a similar approach as in Study 1, with the main difference that participants' language profile (*Group*) was now included as a predictor in the model, in interaction with the *Condition* predictor. We set two *a priori* contrasts for the *Condition* predictor [@schad2020capitalize]: one comparing *Unrelated* and *Related/Non-cognate* trials (sum-coded as `-0.5` and `+0.5`, with *Related/Cognate* trials coded as `0`), and another comparing *Related/Non-cognate* and *Related/Cognate* trials (sum-coded as `-0.5` and `+0.5`, with *Unrelated* trials coded as `0`). In Study 2, the base model $\mathcal{M}_0$ included the the main effects of *Age*, *Condition*, and *Group*, and the two-way interaction between the *Condition* and  *Group* predictors. Contrast coding of the *Condition* predictor was the same as in Study 1. We set one *a priori* contrasts for the *Group* predictor, comparing *Monolingual* with *Bilingual* participants (sum-coded as `-0.5` and `+0.5`, respectively).

@@

```{r coefs}
var_labels <- c(
	"b_Intercept" = "Intercept",
	"b_age_std" = "Age (+1 SD)",
	"b_lp1" = "Group",
	"b_condition1" = "Condition (U vs. RN)",
	"b_condition2" = "Condition (RN vs. RC)",
	"b_condition1:lp1" = "Group \U00D7 Condition (U vs. RN)",
	"b_condition2:lp1" = "Group \U00D7 Condition (RN vs. RC)"
)

model_summary <- get_posterior_summary(model_fits_bcn$fit_0,
									   data_bcn,
									   vars_dict = var_labels,
									   rope_interval = rope_interval
)

@@

Third, it is possible that the modifications of the implicit naming task in the present investigation might have reduce the chances of detecting the anticipated effects. The most critical difference between the original design of the implicit naming task by Mani and Plunkett and that of the present study is the absence of a pre-naming phase during the test phase. Target auditory labels were presented immediately after the offset of the prime picture. It is possible that such time interval was too short for participants to retrieve the phonological label of the prime picture before the target was presented. Further research should investigate the role of prime-target inter-stimulus interval on the time course of lexical access and selection in toddlers.

>>>>

library(ggplot2)
library(ggrepel)
library(gt)
library(patchwork)
library(english)

@@

clrs <- c("#004AAD", "#C8102E", "#FF9E1F")

options(ggplot2.ordinal.fill = clrs,
		ggplot2.ordinal.colour = clrs,
		ggplot2.discrete.fill = clrs,
		ggplot2.discrete.colour = clrs,
		ggplot2.continuous.fill = ggplot2::scale_color_gradient,
		ggplot2.continuous.colour = ggplot2::scale_color_gradient)

@@

In line with previous studies in monolinguals, we further predicted that the strength of the lexical interference effects in the *Related/Non-cognate* and *Related/Cognate* conditions would be associated to participants' vocabulary size. Target word recognition should be delayed by the activation of a larger cohort of phonologically related words [@mani2011phonological; @chow2017spoken; @avila2021longitudinal; @mayor2014infant]. We defined vocabulary size as the amount of words participants were reported to understand in their dominant language by their caregivers. The choice of the dominant language for calculating vocabulary sizes is due to several reasons. First, it allows a more fair comparison between monolinguals (who do not know any language other than their dominant language) and bilinguals (who may know words in a second language). Second, since participants were tested exclusively in their dominant language, their vocabulary size in the dominant language is more likely to be associated to participants' performance in the task. Third, previous work on word recognition in bilinguals suggests that vocabulary size in the dominant language predicts participants' performance better than total vocabulary (in which vocabulary sizes in both languages are summed together) [@marchman2010vocabulary].

@@

	cols_merge(c(age_min, age_max), pattern = "{1}-{2}") |>
	cols_merge(c(age_mean, age_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(doe_catalan_mean, doe_catalan_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(doe_english_mean, doe_english_sd),
			   pattern = "{1} ({2})") |>
	cols_merge(c(doe_spanish_mean, doe_spanish_sd),
			   pattern = "{1} ({2})") |>
	tab_spanner("Sample size", starts_with("n")) |>
	tab_spanner("Age (months)", matches("age")) |>
	tab_spanner("Catalan", matches("catalan")) |>
	tab_spanner("English", matches("english")) |>
	tab_spanner("Spanish", matches("spanish")) |>
	tab_spanner("Degree of Exposure (%)", matches("doe")) |>
	cols_label(n = "Test sessions",
			   n_participants = "Participants",
			   age_mean = "M (SD)",
			   age_min = "Range",
			   doe_catalan_mean = "M (SD)",
			   doe_english_mean = "M (SD)",
			   doe_spanish_mean = "M (SD)") |>
	tab_style(cell_text(style = "italic"),
			  cells_column_labels()) |>
	tab_style(cell_text(style = "italic"),
			  cells_column_spanners())
```

@@

```{r fig-vocabulary}
#| label: fig-vocabulary-oxf
#| fig-cap: "Participant receptive vocabulary sizes across ages and language profiles. Vocabulary size scores from the recurrent participants are shown linked. For visualisation purposes, testing sessions were aggregated at three age points: 21, 25, and 30 months. Points and intervals indicate mean and standard deviations, respectively."
#| fig-width: 5
#| fig-height: 4
#| out-width: 70%
vocab_plot_data <- vocabulary |>
	inner_join(filter(attrition_participants, is_valid_participant),
			   by = join_by(session_id)) |>
	inner_join(select(participants, location, lp, age,
					  child_id, session_id, age_group),
			   by = join_by(child_id, session_id)) |> 
	filter(is_valid_participant) |> 
	select(child_id, session_id, vocab_id, is_imputed, l1_count,
		   lp, location, age_group, age) |> 
	mutate(location = factor(location, 
							 levels = c("Oxford", "Barcelona")),
		   lp = gsub(" \\(English\\)", "", lp),
		   side = case_when(location=="Oxford" ~ "both",
		   				 lp=="Monolingual" ~ "left",
		   				 lp=="Bilingual" ~ "right"),
		   study = if_else(location=="Oxford", 
		   				"Study 1 (English)", "Study 2 (Catalan/Spanish)"))

@@

vocab_plot_data |> 
	filter(study=="Study 1 (English)") |> 
	ggplot(aes(age, l1_count, 
			   side = side)) +
		geom_point(size = 2,
			   shape = 1,
			   colour = "black",
			   alpha = 1/2,
			   stroke = 1) +
	geom_smooth(method = "lm", 
				formula = "y ~ x",
				colour = "black",
				fill = "grey") +
	labs(x = "Age (months)",
		 y = "Vocabulary size",
		 colour = "Group",
		 fill = "Group") +
	scale_y_continuous(limits = c(0, 575),
					   breaks = seq(0, 1e3, 100)) +
	scale_x_continuous(breaks = seq(0, 50, 2)) +
	theme(panel.grid.major.y = element_line(linetype = "dotted",
											colour = "grey"),
		  legend.position = "top",
		  legend.title = element_blank())
```

@@

We created four lists of trials, across which the same target-distractor pair appeared with a different prime, counterbalancing the condition to which it belonged. For instance, in list one the *ball*-*trousers* pair was preceded by *bike* (*Related/Cognate*), by *butterfly* (*Related/Non-cognate*) in list two, and by *star* and *nose* (*Unrelated*) in lists three and four (see Appendix A for a detailed description of the stimuli). @tbl-stimuli shows a detailed summary of the stimuli properties, broken down by trial type and testing language. Trials included in each condition had equivalent length (number of phonemes) and lexical frequency. Lexical frequencies were extracted from the English corpora from the CHILDES database [@macwhinney2000childes; @sanchez2019childes] as counts per million words, and transformed into Zipf scores for easier cross-language comparison [@zipf1945meaning; @van2014subtlex]. In addition to sharing phonological onset, prime and target shared higher overall phonological similarity---as measured using the Levenshtein distance---between their phonological transcriptions---in the *Related/Cognate* and the *Related/Non-cognate* conditions, compared to the *Unrelated* condition. In *Related/Cognate* trials, the prime label in the language of test shared a higher overall phonological similarity with its translation, compared to *Related/Non-cognate* trials. Since the number of cognate and non-cognate primes was balanced in *Unrelated* trials, overall phonological similarity between prime labels prime labels in the *Unrelated* condition had an intermediate overall phonological similarity. Audio durations were also equivalent across conditions.

@@

Testing took place in a sound-proof room at the BabyLab of the University of Oxford. Participants sat on their caregivers' lap in a dimly lit testing booth while the experimenter conducted the experiment from outside. Caregivers were instructed to keep their eyes shut (to avoid recording their gaze, instead of the participant's), to be still, and to avoid interacting with the participant verbally or non-verbally. Participants sat at approximately 65 cm from the eye-tracker and a 23-inches screen with $1929\times1080$ resolution. The study was run on Windows 7 (64-bit), using a custom Matlab script, PresentMate, based on the PsychToolbox-3 extension (3.0.10, 32 bit) [@kleiner2007s; @brainard1997psychophysics; @pelli1997videotoolbox]. Visual fixations were recorded using a Tobii TX300 eye-tracker (Tobii Technology, Stockholm, Sweden) and a 23-in screen of 1920 $\times$ 1080 resolution. The Tobii Analytics SDK 3.0 was used to interact with the eye-tracking while the experiment was running. Sampling rate was set at 120 Hz. A 9-point calibration was performed before every experimental session, in which the picture of a colourful beach ball was presented. We set a 55% grey background for the screen during calibration and stimuli presentation. Auditory stimuli were presented through two loudspeakers located behind the screen, one to each side. The experimenter monitored the experimental from outside the room using a centrally located video camera place above the screen. After a successful calibration the experimenter triggered the onset of the first trial. Trials were presented uninterruptedly and without intervention of the experimenter until the 32 trials were presented, or the experimental session had to be stopped because of the participant's behaviour.

@@

After trials that matched any of the aforementioned exclusion criteria from the data set, we excluded participants who did not provide at least two valid trials in each condition (*n* = `r n_exc_participants`), and participants with a vocabulary size lower than 42, which corresponds to 10% of the words in the OCDI vocabulary checklist (*n* = `r n_exc_participants_vocab`). The final data set included `r format(n_trials_valid, big.mark = ",")` trials from `r n_sessions_valid` testing sessions, generated by `r format(n_participants_valid, big.mark = ",")` distinct participants. Of those participants, `r n_longitudinal[1, 2]` provided data from one experimental session, `r n_longitudinal[2, 2]` provided data from two experimental sessions, and `r n_longitudinal[3, 2]` provided data from three experimental sessions. From the trials included in the final data set, `r n_trials_condition$unrelated_2$n` were *Unrelated* trials (`r n_trials_condition$unrelated$n` previously excluded), were `r n_trials_condition$non_cognate_2$n` *Related/Non-cognate* trials(`r n_trials_condition$non_cognate$n` previously excluded), and `r n_trials_condition$cognate_2$n` were *Related/Cognate* trials (`r n_trials_condition$cognate$n` previously excluded).

@@

r_version <- glue::glue(r_info$R.version$major,
						r_info$R.version$minor,
						.sep = ".")
stan_version <- "2.33.0"

@@

model_summary <- get_posterior_summary(model_fits_oxf$fit_oxf_0,
									   data_bcn,
									   vars_dict = var_labels,
									   rope_interval = rope_interval)

@@

```{r fig-epreds-oxf}
#| label: fig-epreds-oxf
#| fig-cap: "A) Posterior mean predictions of the time course of target fixation in the test phase. B) Posterior mean prediction of the time course of the differences in target looking time between conditions. Intervals represent the 95% CrI of the posterior predictions. Lines indicate the mean of the posterior predictions."
#| fig-width: 6.5
#| fig-height: 6.5
#| out-width: 80%
make_std <- function(x, mean, sd) (x - mean) / sd

@@

plot_epreds + plot_diffs +
	plot_layout(ncol = 1, heights = c(0.6, 0.4)) &
	plot_annotation(tag_levels = "A") &
	# scale_y_continuous(limits = c(0.25, 0.75)) +
	scale_x_continuous(breaks = make_std(seq(0, 17, 3),
										 mean(data_bcn$timebin),
										 sd(data_bcn$timebin)),
					   labels = \(x) format(seq(3e2, 2e3, 3e2),
					   					 big.mark = ",")) &
	theme(legend.position = "top",
		  legend.title = element_blank())
```

@@

We found strong evidence of successful word recognition across participants of all ages, but we did not observe any evidence of phonological priming. English monolingual participants from all ages showed an equivalent pattern of target looking in both Related (Related/Cognate and Related/Non-cognate) and Unrelated trials. In conclusion, we failed to replicate the original studies by @mani2010infant and @mani2011phonological. The absence of a phonological priming effect suggest that either English monolinguals did not generate implicit labels for the prime pictures presented in silence, or that, if generated, such labels did not interact with the subsequent recognition of the target word. Both candidate explanations conflict with both Mani and Plunkett’s studies, and also with previous studies suggesting that infants 12-months and older already generate internal labels when presented with pictures of familiar objects [@duta2012erp; @styles2015infant].

Adding the predictors *Age* or *Vocabulary size* as predictors in the model, in interaction with *Condition* did not increase the fit of the model. This points to neither variable having a substantial influence in participants' target looking behaviour across conditions. These results diverge from previous studies reporting an increment in word recognition speed [@fernald1998rapid; @marchman2008speed], and stronger phonological priming effects in children with larger vocabulary sizes [@mani2011phonological; @chow2017spoken; @avila2021longitudinal]. Overall, this results suggest that our modification of the implicit naming task resulted in the loss of the originally reported effect.

@@

suppressWarnings({
	time_diffs <- participants_bcn |>
		arrange(child_id, session_id, age) |>
		mutate(n_session = 1:n()) |>
		select(child_id, session_id, session_n, age) |>
		filter(child_id %in% participants_bcn$child_id[duplicated(participants_bcn$child_id)]) |>
		pivot_wider(id_cols = child_id,
					names_from = session_n,
					values_from = age,
					names_repair = janitor::make_clean_names) |>
		rowwise() |>
		mutate(diff_1 = x2 - x1, diff_2 = x3 - x2,
			   diff_min = min(diff_1, diff_2, na.rm = TRUE),
			   diff_max = max(diff_1, diff_2, na.rm = TRUE)) |>
		ungroup() |>
		summarise(diff_min = min(diff_min, na.rm = TRUE),
				  diff_max = max(diff_max, na.rm = TRUE))
})
```

@@

We collected vocabulary data using parental responses to the Barcelona Vocabulary Questionnaire [BVQ, @garcia-castro2023bvq], an online vocabulary checklist developed to assess the vocabulary size of Catalan-Spanish bilingual toddlers, and inspired in several adaptations of the the Communicative Developmental Inventory [CDI, @fenson1994variability]. This questionnaire has four versions, each including a different but overlapping subset of words, from a total pool of `r n_pool` words from `r n_categories` functional-semantic categories. Each version included a Catalan and a Spanish vocabulary checklist. Catalan checklists contained between `r n_items$min[1]` and `r n_items$min[2]` words, and Spanish checklists contained between `r n_items$min[2]` and `r n_items$min[2]` words (see the Methods section of Chapter 1 for a detailed description of the questionnaire. Participants were randomly allocated to one of the four versions. Recurrent participants were always allocated to the same version. Families received a link to the BVQ immediately after each experimental session, and were given two weeks to fill it.

@@

```{r fig-vocabulary}
#| label: fig-vocabulary-bcn
#| fig-cap: "Participant receptive vocabulary sizes across ages and language profiles. Points and intervals indicate mean and standard deviations, respectively."
#| fig-width: 6
#| fig-height: 3
vocab_plot_data |> 
	mutate(lp = factor(lp, levels = c("Monolingual", "Bilingual"))) |> 
	filter(study=="Study 2 (Catalan/Spanish)") |> 
	ggplot(aes(age, l1_count)) +
	facet_wrap(~lp) +
	geom_point(size = 2,
			   shape = 1,
			   colour = "black",
			   alpha = 1/2,
			   stroke = 1) +
	geom_smooth(method = "lm", 
				formula = "y ~ x",
				colour = "black",
				fill = "grey") +
	labs(x = "Age (months)",
		 y = "Vocabulary size",
		 colour = "Group",
		 fill = "Group") +
	scale_x_continuous(breaks = seq(0, 40, 2)) +
	scale_y_continuous(limits = c(0, 575),
					   breaks = seq(0, 1e3, 100)) +
	scale_x_continuous(breaks = seq(0, 50, 2)) +
	theme(panel.grid.major.y = element_line(linetype = "dotted",
											colour = "grey"),
		  legend.position = "none",
		  axis.title.x = element_blank(),
		  legend.title = element_blank())
```

@@

Same as in Study 1. We run the study on Windows 10 64-bit, using custom Matlab (2018a 64-bit) script using the PsychToolbox-3 extension (3.0.15 64-bit) to present the stimuli on a 23-inches screen with $1929\times1080$ resolution, and the Tobii Analytics SDK 3.0 to interact with the eye-tracker (Tobii TX300 and Tobii Pro Sprectrum, Tobii Technology, Stockholm, Sweden) while the experiment was running.

@@

```{r data set}
attrition_bcn <- filter(attrition, location=="Barcelona")
n_trials <- nrow(attrition_bcn)
n_sessions <- n_distinct(attrition_bcn$session_id)
n_participants <- n_distinct(attrition_bcn$child_id)
n_trials_valid <- nrow(filter(
	attrition_bcn,
	is_valid_participant,
	is_valid_trial
))
n_sessions_valid <- sum(distinct(attrition_bcn,
								 session_id,
								 .keep_all = TRUE)$is_valid_participant)
n_participants_valid <- n_distinct(
	attrition_bcn[attrition_bcn$is_valid_participant, ]$child_id)

n_exc_prime <- sum(!attrition_bcn$is_valid_gaze_prime)
n_exc_test <- sum(!attrition_bcn$is_valid_gaze_test)
n_exc_test_each <- sum(!attrition_bcn$is_valid_gaze_test_each)
n_exc_participants <- sum(!distinct(attrition_bcn,
									session_id,
									.keep_all = TRUE)$is_valid_participant)
n_exc_participants_vocab <- sum(!distinct(attrition_bcn,
										  session_id,
										  .keep_all = TRUE)$is_valid_vocab_size)

n_longitudinal <- attrition_bcn |>
	filter(is_valid_participant,
		   is_valid_trial) |>
	distinct(child_id, session_n) |>
	count(child_id, name = "times") |>
	count(times)
```

@@

We modelled the data following a similar approach as in Study 1, with the main difference that participants' language profile (*Group*) was now included as a predictor in the model, in interaction with the *Condition* predictor. We set two *a priori* contrasts for the *Condition* predictor [@schad2020capitalize]: one comparing *Unrelated* and *Related/Non-cognate* trials (sum-coded as `-0.5` and `+0.5`, with *Related/Cognate* trials coded as `0`), and another comparing *Related/Non-cognate* and *Related/Cognate* trials (sum-coded as `-0.5` and `+0.5`, with *Unrelated* trials coded as `0`). In Study 2, the base model $\mathcal{M}_0$ included the the main effects of *Age*, *Condition*, and *Group*, and the two-way interaction between the *Condition* and  *Group* predictors. Contrast coding of the *Condition* predictor was the same as in Study 1. We set one *a priori* contrasts for the *Group* predictor, comparing *Monolingual* with *Bilingual* participants (sum-coded as `-0.5` and `+0.5`, respectively).

@@

```{r coefs}
var_labels <- c(
	"b_Intercept" = "Intercept",
	"b_age_std" = "Age (+1 SD)",
	"b_lp1" = "Group",
	"b_condition1" = "Condition (U vs. RN)",
	"b_condition2" = "Condition (RN vs. RC)",
	"b_condition1:lp1" = "Group \U00D7 Condition (U vs. RN)",
	"b_condition2:lp1" = "Group \U00D7 Condition (RN vs. RC)")

model_summary <- get_posterior_summary(model_fits_bcn$fit_0,
									   data_bcn,
									   vars_dict = var_labels,
									   rope_interval = rope_interval)

@@

Third, and most likely, the modifications of the implicit naming task in the present investigation might have reduce the chances of detecting the anticipated effects. The most critical difference between the original design of the implicit naming task by Mani and Plunkett and that of the present study is the absence of a pre-naming phase during the test phase. Target auditory labels were presented immediately after the offset of the prime picture. It is possible that such time interval was too short for participants to retrieve the phonological label of the prime picture before the target was presented. Such failure to generate phonological word-forms for prime labels would have prevented participants in Studies 1 and 2 from being affected by phonological priming effects during target word recognition.
